Level-Set method on \texttt{p4est}

### 2014-12-02 paper/tex/scaling_results.tex
Updated scaling results

\subsection{Semi-Lagrangian}
To test the scalability of the Semi-Lagrangian algorithm \ref{alg:semi-lagrangian} we consider a slightly modified version of the Enright's rotation test \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}, i.e. we advect a sphere of radius 0.35 located at $(0.4, 0.4, 0.4)$ with a divergent free velocity field given by:
\bean
u(x,y,z) &=& 2\sin(\pi x)^2 \sin(2\pi y) \sin(2 \pi z), \\
v(x,y,z) &=& -\sin(\pi y)^2 \sin(2\pi x) \sin(2 \pi z),\\
w(x,y,z) &=& -\sin(\pi z)^2 \sin(2\pi x) \sin(2 \pi y). 
\eean
To understand the effect of CFLnumber on the scalability of the algorithm we perform one step of the Semi-Lagrangian algorithm for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$. We also perform the test for two set of initial girds, a small grid with $l_\max = 10$ and a larger grid with $l_\max = 12$. After one advection step, each of these grids have approximately 15M and 255M nodes, respectively. 
\begin{table}[htbp]
	\begin{minipage}{.5\linewidth}
		\begin{center}
		\begin{tabular}{ccccccc}
			\hline
				CFL & 16 & 32 & 64 & 128 & 256 & 512 \\
				\hline
				1   & 2 & 2 & 3 & 3 & 3 & 3 \\
				10  & 3 & 3 & 3 & 3 & 3 & 3 \\
				100 & 6 & 6 & 6 & 6 & 6 & 6 \\
			\hline
		\end{tabular}
		\caption*{(a) $l_\max = 10$}
		\end{center}
	\end{minipage}%
	\begin{minipage}{.5\linewidth}
		\begin{center}
		\begin{tabular}{ccccccc}
			\hline
				CFL & 128 & 256 & 512 & 1024 & 2048 & 4096 \\
			\hline
				1   & 3 & 3 & 3 & 3 & 3 & 3 \\
				10  & 3 & 3 & 4 & 4 & 4 & 4 \\
				100 & 6 & 6 & 6 & 6 & 6 & 7 \\
			\hline
		\end{tabular}
		\caption*{(b) $l_\max = 12$}
		\end{center}
	\end{minipage}	
	\caption{Number of iterations required for grid construction in algorithm \ref{alg:semi-lagrangian} for the rotation test on a (a) level-10 and (b) level-12 octree with approximately 15M and 255M nodes, respectively. Note how the iteration count increases with the CFL number but is almost independent of number of processors. The slight dependence of iteration count on number of processors is most likely due to dependence of round-off errors on the number of processors. Nonetheless, close examination of generated octrees (data not shown) reveals that they are identical and independent of the number of processors used to perform the test.}
	\label{tab:semilagrangian}
\end{table}

@@

An increase in the CFL number has two effects on the algorithm. First, a larger fraction of departure points land in domains of remote processors. Moreover, these points are potentially dispersed across a larger number of processors. This means both the number MPI messages and their volume increases for larger CFL numbers. Second, as more points are shipped to remote processors for interpolation, there is a greater chance that interpolation load is unbalanced across processors. This is specially true for regions of space in which streamlines cluster. Both factors contribute to decrease the scalability of the algorithm at larger CFL numbers. To shed more light on this, we have recorded a complete history of communication pattern in the interpolation step.

\begin{figure}[htbp]
	\begin{center}
		\subfigure[Semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_1_SemiLagrangian.pdf}}
		\subfigure[Semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_10_SemiLagrangian.pdf}}
		\subfigure[Semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_100_SemiLagrangian.pdf}}
		\\
		\subfigure[Interpolation, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_1_Interpolation.pdf}}
		\subfigure[Interpolation, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_10_Interpolation.pdf}}
		\subfigure[Interpolation, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_100_Interpolation.pdf}}
	\end{center}
	\caption{Strong scaling of a single iteration of algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-10 octree with approximately 13M cells and 15M nodes. Top row: scaling of various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$. Bottom row: breakdown of various components of the interpolation phase for the same CFL numbers. Note that the maximum time has been scaled with the number of iterations required to build the tree (c.f. table \ref{tab:semilagrangian}).}
	\label{fig:semilagrangian_small}
\end{figure}

@@

@@

% \subsection{Poisson solver}

>>>>

\subsection{Semi-Lagrangian}
To test the scalability of the Semi-Lagrangian algorithm \ref{alg:semi-lagrangian} we consider a slightly modified version of the Enright's rotation test \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}, i.e. we advect a sphere of radius 0.35 located at $(0.4, 0.4, 0.4)$ with a divergent free velocity field given by:
\bean
u(x,y,z) &=& 2\sin(\pi x)^2 \sin(2\pi y) \sin(2 \pi z), \\
v(x,y,z) &=& -\sin(\pi y)^2 \sin(2\pi x) \sin(2 \pi z),\\
w(x,y,z) &=& -\sin(\pi z)^2 \sin(2\pi x) \sin(2 \pi y). 
\eean
To understand the effect of CFLnumber on the scalability of the algorithm we perform one step of the Semi-Lagrangian algorithm for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$. We also perform the test for two set of initial girds, a small grid with $l_\text{max} = 10$ and a larger grid with $l_\text{max} = 12$. After one advection step, each of these grids have approximately 15M and 255M nodes, respectively. 
\begin{table}[htbp]
	\begin{minipage}{.5\linewidth}
		\begin{center}
		\begin{tabular}{ccccccc}
			\hline
				CFL & 16 & 32 & 64 & 128 & 256 & 512 \\
				\hline
				1   & 2 & 2 & 3 & 3 & 3 & 3 \\
				10  & 3 & 3 & 3 & 3 & 3 & 3 \\
				100 & 6 & 6 & 6 & 6 & 6 & 6 \\
			\hline
		\end{tabular}
		\caption*{(a) $l_\text{max} = 10$}
		\end{center}
	\end{minipage}%
	\begin{minipage}{.5\linewidth}
		\begin{center}
		\begin{tabular}{ccccccc}
			\hline
				CFL & 128 & 256 & 512 & 1024 & 2048 & 4096 \\
			\hline
				1   & 3 & 3 & 3 & 3 & 3 & 3 \\
				10  & 3 & 3 & 4 & 4 & 4 & 4 \\
				100 & 6 & 6 & 6 & 6 & 6 & 7 \\
			\hline
		\end{tabular}
		\caption*{(b) $l_\text{max} = 12$}
		\end{center}
	\end{minipage}	
	\caption{Number of iterations required for grid construction in algorithm \ref{alg:semi-lagrangian} for the rotation test on a (a) level-10 and (b) level-12 octree with approximately 15M and 255M nodes, respectively. Note how the iteration count increases with the CFL number but is almost independent of number of processors. The slight dependence of iteration count on number of processors is most likely due to dependence of round-off errors on the number of processors. Nonetheless, close examination of generated octrees (data not shown) reveals that they are identical and independent of the number of processors used to perform the test.}
	\label{tab:semilagrangian}
\end{table}

@@

An increase in the CFL number has two effects on the algorithm. First, a larger fraction of departure points land in domains of remote processors. Moreover, these points are potentially dispersed across a larger number of processors. This means both the number MPI messages and their volume increases for larger CFL numbers. Second, as more points are shipped to remote processors for interpolation, there is a greater chance that interpolation load is unbalanced across processors. This is specially true for regions of space in which streamlines cluster. Both factors contribute to decrease the scalability of the algorithm at larger CFL numbers. To shed more light on this, we have recorded a complete history of communication pattern in the interpolation step. Figures \ref{fig:communication_128} and \ref{fig:communication_4096} illustrates the effect of CFL number on three different metrics for $p= 128$ and $p=4096$ processors in the first and last iterations of the Semi-Lagrangian algorithm. 
\begin{figure}[htbp]
	\begin{center}
		\subfigure[Semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_1_SemiLagrangian.pdf}}
		\subfigure[Semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_10_SemiLagrangian.pdf}}
		\subfigure[Semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_100_SemiLagrangian.pdf}}
		\\
		\subfigure[Interpolation, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_1_Interpolation.pdf}}
		\subfigure[Interpolation, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_10_Interpolation.pdf}}
		\subfigure[Interpolation, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_100_Interpolation.pdf}}
	\end{center}
	\caption{Strong scaling of a single iteration of algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-10 octree with approximately 13M cells and 15M nodes. Top row: scaling of various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$. Bottom row: breakdown of various components of the interpolation phase for the same CFL numbers. Note that the maximum time has been scaled with the number of iterations required to build the tree (c.f. table \ref{tab:semilagrangian}).}
	\label{fig:semilagrangian_small}
\end{figure}

@@

\begin{figure}[htbp]
	\begin{center}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/point_first_128.pdf}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/message_first_128.pdf}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/volume_first_128.pdf}}
		\\
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/point_last_128.pdf}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/message_last_128.pdf}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/volume_last_128.pdf}}		
	\end{center}
	\caption{Performance indicators of the first (top row) and last (bottom row) iterations of the Semi-Lagrangian algorithm for the advection test for $p=128$ processors. Increasing the CFL number greatly causes load imbalance in the amount of computation done during interpolation (a) and increases the communication volume (c). However the CFL number does not seem to appreciably affect the number of messages sent by processors (b). During the last Semi-Lagrangian iteration the initial grid, $G_0$ (c.f. algorithm \ref{alg:semi-lagrangian}), is very close to the final grid. As a result load imbalance is considerably improved (d). Curiously, however, communication pattern does not seem to be affected by much between first and last iterations (e-f).}
	\label{fig:communication_128}
\end{figure}

\begin{figure}[htbp]
	\begin{center}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/point_first_4096.pdf}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/message_first_4096.pdf}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/volume_first_4096.pdf}}
		\\
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/point_last_4096.pdf}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/message_last_4096.pdf}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/volume_last_4096.pdf}}		
	\end{center}
	\caption{Performance indicators of the first (top row) and last (bottom row) iterations of the Semi-Lagrangian algorithm for the advection test for $p=4096$ processors. Increasing the CFL number greatly causes load imbalance in the amount of computation done during interpolation (a) and increases the communication volume (c). However the CFL number does not seem to appreciably affect the number of messages sent by processors (b). During the last Semi-Lagrangian iteration the initial grid, $G_0$ (c.f. algorithm \ref{alg:semi-lagrangian}), is very close to the final grid. As a result load imbalance is considerably improved (d). Curiously, however, communication pattern does not seem to be affected by much between first and last iterations (e-f).}
	\label{fig:communication_4096}
\end{figure}

@@

% \subsection{Poisson solver}
% p = 128, folder =  host/large/100, mode = _first_ (min, max, avg, std) 
% point  : 1.56e+06, 2.82e+06, 2.07e+06, 1.65e+05
% message: 19, 131, 41.78, 19.73
% volume : 5.81e+00, 3.45e+01, 1.97e+01, 6.46e+00

% p = 128, folder =  host/large/10, mode = _first_ (min, max, avg, std) 
% point  : 1.98e+06, 2.21e+06, 2.07e+06, 3.52e+04
% message: 22, 131, 42.92, 19.56
% volume : 2.58e+00, 9.64e+00, 4.92e+00, 1.24e+00

% p = 128, folder =  host/large/1, mode = _first_ (min, max, avg, std) 
% point  : 2.02e+06, 2.20e+06, 2.07e+06, 2.97e+04
% message: 22, 131, 42.97, 19.53
% volume : 2.45e+00, 9.03e+00, 4.21e+00, 1.17e+00

% p = 4096, folder =  host/large/100, mode = _first_ (min, max, avg, std) 
% point  : 8.28e+02, 4.98e+05, 7.57e+04, 3.14e+04
% message: 11, 373, 41.18, 30.85
% volume : 5.28e-01, 1.01e+01, 2.65e+00, 9.24e-01

% p = 4096, folder =  host/large/10, mode = _first_ (min, max, avg, std) 
% point  : 5.56e+04, 1.63e+05, 7.57e+04, 6.07e+03
% message: 17, 387, 46.73, 31.78
% volume : 4.01e-01, 3.06e+00, 8.40e-01, 2.21e-01

% p = 4096, folder =  host/large/1, mode = _first_ (min, max, avg, std) 
% point  : 6.67e+04, 1.59e+05, 7.57e+04, 4.86e+03
% message: 18, 387, 47.55, 31.72
% volume : 4.01e-01, 2.93e+00, 7.25e-01, 1.95e-01

% p = 128, folder =  host/large/100, mode = _last_ (min, max, avg, std) 
% point  : 1.96e+06, 2.17e+06, 2.06e+06, 3.37e+04
% message: 22, 133, 41.86, 18.62
% volume : 1.04e+01, 6.63e+01, 2.98e+01, 1.16e+01

% p = 128, folder =  host/large/10, mode = _last_ (min, max, avg, std) 
% point  : 2.02e+06, 2.16e+06, 2.07e+06, 2.76e+04
% message: 22, 131, 42.88, 19.53
% volume : 3.22e+00, 1.26e+01, 6.24e+00, 1.43e+00

% p = 128, folder =  host/large/1, mode = _last_ (min, max, avg, std) 
% point  : 2.02e+06, 2.19e+06, 2.07e+06, 2.93e+04
% message: 22, 131, 42.89, 19.54
% volume : 2.72e+00, 9.54e+00, 4.46e+00, 1.12e+00

% p = 4096, folder =  host/large/100, mode = _last_ (min, max, avg, std) 
% point  : 6.30e+04, 1.10e+05, 7.57e+04, 4.20e+03
% message: 16, 357, 41.77, 22.66
% volume : 9.76e-01, 4.86e+00, 3.78e+00, 5.69e-01

% p = 4096, folder =  host/large/10, mode = _last_ (min, max, avg, std) 
% point  : 6.65e+04, 1.33e+05, 7.57e+04, 4.50e+03
% message: 19, 393, 46.68, 27.93
% volume : 4.40e-01, 4.80e+00, 2.14e+00, 9.88e-01

% p = 4096, folder =  host/large/1, mode = _last_ (min, max, avg, std) 
% point  : 6.69e+04, 1.55e+05, 7.57e+04, 4.73e+03
% message: 18, 392, 47.55, 31.19
% volume : 4.13e-01, 3.91e+00, 9.68e-01, 2.62e-01
